---
title: "Day 3 Homework"
author: "Katti McNally"
date: "Thursday, January 08, 2015"
output: html_document
---

```{r}
library(reshape2)
install.packages("dplyr")
library(dplyr)
```

#using updated dataset

*Read data in*
```{r}
candidate <- read.csv(file = "candidate.csv", header = TRUE, sep = ",")
head(candidate)
colnames(candidate)
```

# 1) Long and Wide Versions of Data

## Long Data
```{r}
melt <- melt(candidate,id.vars=c("cycle", "District", "Name"),measure.vars = c( "total.disbursements", "total.receipts", "total.indiv.contrib", "total.pac.contribs"), variable.name = "campaignFinance", value.name = "Dollars", na.rm = TRUE)
```

##Wide Data
```{r}
cast <- dcast(melt,cycle+District+Name~campaignFinance,value.var ="Dollars")
```

- casting with a lot of variables
```{r eval = FALSE}

mydata.wide <- dcast(mydata.long, [80 columns here] ~ split) #split is representing the column you wish to split

columns.keep <- colnames(mydata.long)[1:80]
columns.keep.pasted <- paste(columns.keep, collapse = " + ")
my.cast.formula <- as.formula(paste(columns.keep.pasted, "~split"))
class(my.cast.formula) # "formula"

mydata.wide <- dcast(mydata.long, formula = my.cast.formula)

#working with "everything but"
sleep[, -1] #leave out the first column
sleep[, -c(1,2)] #leave out the first two columns
#this shit also has to be in numbers, not in strings

```


# 2) Transforming a variable

## Taking the log of the Dollars column in the melted version
```{r}
melt$logDollars <- log(melt$Dollars)
summary(melt)
```

# 3) Recoding a factor

## Recoding Party from 100 and 200 to Democrat and Republican
```{r}
levels(candidate$Party) <- c("Democrat", "Republican")
levels(candidate$Party)
```

# 4) New Factor with values conditional on another variable
```{r}
as.factor(candidate$cand.gender)
candidate$Gender <- "Female" 
candidate$Gender[candidate$cand.gender != 1] <- "Male"  
head(candidate)
```

# 5) Reorder factor levels
```{r}
candidate$Gender <- as.factor(candidate$Gender)
levels(candidate$Gender)
candidate$Gender <- relevel(candidate$Gender, "Male")
levels(candidate$Gender)
```

# 6) Recode factor as multiple dimensions

## Change names to first and last name
```{r}
cand.name <- as.factor(candidate$Name)
candidate.2 <- cbind(candidate, 
                     colsplit(candidate$Name, pattern = ",", 
                              names = c("lastName", "firstName")))
head(candidate.2)
```

other weird way to do this shit...
```{r eval = FALSE}
mydata$var <- ifelse(grepl("left", mydata$var)"back", "front")
#or something like this...
```

# 7) Aggregate variable into cell means using `dcast()`
```{r}
cast <- dcast(melt,cycle+District~campaignFinance,value.var ="Dollars",fun.aggregate=mean, na.rm = TRUE)
```
