---
title: "Day 4 Homework"
output: html_document
---

##Katti McNally, HW4

```{r}
library(reshape2)
```

1. Change the column names of your data to something easier to work with.  If you like your column names, change them into something else reasonable, just for practice.

```{r}
candidate <- read.csv("candidate.csv")
cand.oldnames <- colnames(candidate)
colnames(candidate)[2] <- "name"
colnames(candidate)[3] <- "district"
colnames(candidate)[4] <- "experience"
colnames(candidate)[5] <- "party"
colnames(candidate)[6] <- "per.wom.stateleg"
colnames(candidate)[9] <- "total.disburse"
colnames(candidate)[10] <- "total.receipt"
colnames(candidate)[12] <- "total.pac.contrib"
colnames(candidate)[13] <- "primary.win"
colnames(candidate)[14] <- "gen.win"
cand.newnames <- colnames(candidate)
data.frame(cand.oldnames, cand.newnames)
```

2. List ALL the issues you can see where it looks like your data is being read in a weird way.  If your data looks perfect, find some way to mess it up :-)

```{r}
summary(candidate)
class(candidate$primary.win)
class(candidate$gen.win)
class(candidate$gen.elect.pct)
class(candidate$cand.gender) #should be factor
class(candidate$party) #should be factor, relabel 
class(candidate$name) #should be character
class(candidate$experience) #should be factor, max should be 1, not 11
class(candidate$district)
```

3. Pick one or two of the most egregious or problematic of the issues in #2 and fix them.

```{r}
candidate$name <- as.character(candidate$name)

candidate$experience <- as.factor(candidate$experience)
candidate$experience[candidate$experience == 11] <- 1
summary(candidate$experience)
levels(candidate$experience)
candidate.1 <- subset(candidate, experience != "11")
summary(candidate.1$experience)
candidate.2 <- droplevels(candidate.1)
summary(candidate.2$experience)
summary(candidate.2)

candidate <- candidate.2
summary(candidate)
```

4. Check for NAs, and report where you see them (try to pinpoint where they are, not just "5 NAs in column seven".

```{r}
summary(candidate)
#there are 4 NAs in total.disburse, total.receipt, total.indiv.contrib, and total.pac.distrib.  Presumable these are in the same four rows, but should check just in case...
candidate.a <- candidate[!is.na(candidate$total.disburse), ]
summary(candidate.a)
#all of the NAs for those four variables are in the same row
#there are also 245 NAs in gen.elect.pct.  These are all the observations for candidates that ran in the primary, but not in the general

#checking which NAs...
#which(is.na(mydata[, 1]))

#another way--returns TRUE and FALSE vector for every row of the data that has no NAs
#complete.cases(mydata)
#mydata[!complete.cases(mydata), ]
#which(!complete.cases(mydata))
```

5. Decide what you think you should do about the NAs, and say why.  If you can, try to implement this decision.

```{r}
#eliminate the four rows with NAs for the financials
candidate <- candidate.a
summary(candidate)
#it is not surprising that gen.elect.pct has so many NAs, since not all of the candidates in the dataset ran in the general election.  I'm going to make a subset of the data that only has the candidates with values for gen.elect.pct, rather than just getting rid of the column
gen.candidate <- candidate[!is.na(candidate$gen.elect.pct), ]
summary(gen.candidate)
```

6. Remove any problematic rows and/or columns.  Say why you want to remove them.  If there aren't any problems, describe why you think there are no problems.

- I don't think are any rows are columns that need to be removed, now that the NAs have been taken care of.  The rest of them look as they should.

7. Re-code at least one factor.  Provide a table that shows the correspondence between old levels and new levels.

```{r}
class(candidate$party)
candidate$party <- as.factor(candidate$party)
oldlevels <- levels(candidate$party)
oldlevels
levels(candidate$party) <- c("Democrat", "Republican")
levels(candidate$party)
newlevels <- levels(candidate$party)
newlevels
data.frame(oldlevels, newlevels)
```

8. Run TWO DIFFERENT simple analyses or statistical tests, such as linear regression (`lm()`), logistic regression (`glm()`), correlation test (`cor.test()`), t-test (`t.test()`), or non-parametric tests (e.g., `wilcox.test()`).  For each of these:
  - Describe why you are doing this analysis, i.e., what question is it answering?
  - I won't judge you on statistical expertise!  (though I will make comments if I think I can be helpful)
  - Report some key statistics from the analysis, using inline code

- (1) t-test to evaluate if female republican candidates have a ideology score (cfscore) that is statistically different from male republican candidates on average
```{r}
female <- candidate$cfscore[candidate$cand.gender == 1 & candidate$party == "Republican"]
male <- candidate$cfscore[candidate$cand.gender == 0 & candidate$party == "Republican"]

t.test(female, male)

ttest.results <- t.test(female, male)
summary(ttest.results)

names(ttest.results)
est.female <- ttest.results$estimate[1]
est.male <- ttest.results$estimate[2]
p.value <- ttest.results$p.value
t.stat <- ttest.results$statistic
```
  - The mean cfscore for female Republican candidates was `r round(est.female, 4)`, while the cfscore for male Republican candidates was `r round(est.male, 4)`.  These two values are not statistically different from one another, as indicated by the small t statistic of `r round(t.stat, 4)` and the high p value of `r round(p.value, 4)`.


- (2) logistic regression to determine if male Republican candidates are more likely to win their primary elections
```{r}
logit.results <- glm(formula = primary.win ~ experience + per.wom.stateleg + cfscore + cand.gender + total.receipt, data = candidate, subset = party == "Republican",  family = "binomial")

summary(logit.results)
logit.summary <- summary(logit.results)

names(logit.summary)

est.female <- logit.summary$coefficients["cand.gender", "Estimate"]
st.err.female <- logit.summary$coefficients["cand.gender", "Std. Error"]
p.value.female <- logit.summary$coefficients["cand.gender", "Pr(>|z|)"]
```
  - Female Republican candidates are less likely to win their primary election, as shown by the estimate of `r round(est.female, 4)`.  However, this is just outside of conventional levels of statistical significance with a p value of `r round(p.value.female, 2)` and a standard error of `r round(st.err.female, 4)`.