---
title: "Day 8 Homework"
output: html_document
---

1. Try to recreate the results of Homework 6 (or your intended results) using the `dplyr` functions.

```{r}
library(dplyr)
library(reshape2)

candidate <- read.csv("candidate.csv")

#average cfscore for each district
candidate %>%
  filter(Party == 200) %>%
  group_by(District) %>%
  summarize(mean.dist = mean(cfscore))
```

2. Find another kind of analysis/transformation/summary stat that you want to apply across multiple groups in your data. Use the functions from `reshape2` and `dplyr` to get these results.

```{r}
#min and max disbursements for each district
candidate %>%
  melt(measure.var = 9:12, value.name = "dollars", variable.name = "Financials") %>%
  group_by(District, Financials) %>%
  summarize(min = min(dollars), max = max(dollars))
```

3. For more practice, try another example of something like #2 (find another stat or grouping or something).

```{r}
candidate %>%
  melt(measure.var = 9:12, value.name = "dollars", variable.name = "Financials") %>%
  filter(!is.na(dollars)) %>%
  group_by(cand.gender, Financials) %>%
  summarize(mean = mean(dollars))
```

4. Do yet another case of #2 above, but this time throw in a transformation, too.

```{r}
candidate %>%
  melt(measure.var = 9:12, value.name = "dollars", variable.name = "Financials") %>%
  mutate(log.dollars = log(dollars)) %>%
  filter(cfscore > sd(abs(cfscore))) %>%
  group_by(District, Financials) %>%
  summarize(minlog = min(log.dollars), maxlog = max(log.dollars))
```

