---
title: "Day 9 Homework"
output: html_document
---
  Katti McNally
  
```{r}
library(dplyr)
library(reshape2)

candidate <- read.csv("candidate.csv")
```

1. Write a function to do a bunch of clean-up and formatting to your data.

```{r}
cleanup <- function(data) {
  clean <- { candidate[!is.na(candidate$total.disburse), ] 
    levels(candidate$Party) <- c("Democrat", "Republican") 
    melt(candidate, measure.vars = c("total.disbursements", "total.receipts", "total.indiv.contrib", "total.pac.contribs")) 
    cbind(candidate, colsplit(as.factor(candidate$Name), pattern = ",", names = c("lastName", "firstName")))
    }
print(summary(clean))
}
```

2. Make sure this function works on your data.

```{r}
cleanup(candidate)
```

3. Get some subset of your data, or find a similar-but-not-identical data set.  Change up the column order.  Re-run your function on this new data.  Does it work?

```{r}
candidate.1 <- candidate[1:100, c(9:12, 1:8, 13)]
summary(candidate.1)

cleanup(candidate.1)
```

4. Fix your function to be able to work on the new data.

```{r}
cleanup <- function(data) {
  clean <- { data[!is.na(data$total.disburse), ] 
    levels(data$Party) <- c("Democrat", "Republican") 
    melt(data, measure.vars = c("total.disbursements", "total.receipts", "total.indiv.contrib", "total.pac.contribs")) 
    cbind(data, colsplit(as.factor(data$Name), pattern = ",", names = c("lastName", "firstName")))
    }
print(summary(clean))
}

cleanup(candidate.1)
```

5. Write a function to do some kind of summary stat, and use the `dplyr` functions to run this function over groups of your data.

```{r}
means <- function(data) {
  result <- {
  data %>%
  melt(measure.var = 9:12, value.name = "dollars", variable.name = "Financials") %>%
  filter(!is.na(dollars)) %>%
  summarize(mean = mean(dollars))
  }
result
}

means(candidate)
means(candidate.1)
```

## Additional stuff

```{r eval = FALSE}

myfunc <- function(data = mydata, column = "variable")
  
mydata[, column] #"$" doesn't as well to identify columns in a function...

x <- 1:10
"x"
get("x") #this lets you recover the object with this name (the object represented by the string "x")

miris <- iris
myfunc <- function(data, colname){
  library(dplyr)
  output <- filter(data, data[, colname] == "setosa")
  output
}

```

